Curve counting in genus one: elliptic singularities {\it \&} relative geometry

### 2019-06-28 EllipticSingularitiesSplitting.tex
minor changes

This reconstruction result is most directly a generalization of Gathmann's calculation~\cite{Ga}. A higher genus reconstruction result for the ordinary theory is proved in a paper of Maulik and Pandharipande~\cite{MaulikPandharipande}, but is of a fundamentally different approach than ours. In particular, localization is never used in the analysis here, nor in the work of Gathmann.

@@

We achieve the first step using techniques that are entirely within tropical and logarithmic geometry, substantially simplifying Gathmann's calculation. For the second step, we contend with the interaction of the relative splitting formula and the geometry of elliptic singularities, which is entirely new (\S \ref{subsection C0 splitting}). We express the factorisation condition -- central to the constructions in~\cite{RSPW,RSPW2} -- in terms of tautological classes, leading to Theorem~\ref{thm: recursion}.

\subsection{Broader contributions} Beyond genus one, we pass through some useful general techniques. First, we explain how to describe degenerate moduli spaces as fibre products. This is substantially more delicate than the standard situation for two reasons. On the one hand, we need to study the interaction of the relative geometry with elliptic singularities, which involves a study of the factorization condition in terms of tautological vector bundles. On the other hand, the degenerate strata of moduli spaces are not simply fiber products of spaces with smaller numerical invariants, but compactified torus bundles over these strata, due to the introduction of logarithmic blowups. This geometry is present throughout logarithmic enumerative geometry, particularly in the degeneration formula, and we provide a first systematic procedure that explains how to use tropical geometry to calculate with these torus bundles~\cite{R19}. In addition, the results of~\cite{RSPW} suggest the existence of a reduced higher genus Gromov--Witten theory formed by replacing contracted elliptic components with singularities\footnote{We have been informed by Jonathan Wise that his Ph.D. candidate Sebastian Bozlee has constructed these higher genus reduced Gromov--Witten invariants. A manuscript is in preparation.}. The discussion here is likely to carry over to this setting.

@@

We have produced a factorization
\begin{equation*}\widetilde\Dcal\to\widetilde \Ecal\to\Ecal,
\end{equation*}
where the first map is flat and the second map is a logarithmic modification. $\widetilde\Dcal$ is a compactification of a torus bundle over $\widetilde\Ecal$ whose strata can be read off from the tropicalization; in particular, we have the information of how (polynomials in) the classes of boundary strata push forward to $\widetilde\Ecal$. In fact, we can say more and describe the generic fibre of $\widetilde\Dcal \to \widetilde\Ecal$:
\begin{lemma}\label{lem:generic_proj_bundle}
Generically over $\Ecal$, $\widetilde \Dcal\to\widetilde\Ecal$ is the projectivisation of a tautological vector bundle.
\end{lemma}

>>>>

This reconstruction result is most directly a generalization of Gathmann's calculation~\cite{Ga}. A higher genus reconstruction result for the ordinary theory is proved in a paper of Maulik and Pandharipande~\cite{MaulikPandharipande}, but is of a fundamentally different approach than ours. In particular, localisation is never used in the analysis here, nor in the work of Gathmann.

@@

For the second step, we contend with the interaction of the relative splitting formula and the geometry of elliptic singularities, which is entirely new (\S \ref{subsection C0 splitting}). We express the factorisation condition -- central to the constructions in~\cite{RSPW,RSPW2} -- in terms of tautological classes, leading to Theorem~\ref{thm: recursion}.

\subsection{Broader contributions} Beyond genus one, we pass through some useful general techniques. First, we explain how to describe degenerate moduli spaces as fibre products. This is substantially more delicate than the standard situation for two reasons. On the one hand, we need to study the interaction of the relative geometry with elliptic singularities, which involves a study of the factorisation condition in terms of tautological vector bundles. On the other hand, the degenerate strata of moduli spaces are not simply fiber products of spaces with smaller numerical invariants, but compactified torus bundles over these strata, due to the introduction of logarithmic blowups. This geometry is present throughout logarithmic enumerative geometry, particularly in the degeneration formula, and we provide a first systematic procedure that explains how to use tropical geometry to calculate with these torus bundles~\cite{R19}. In addition, the results of~\cite{RSPW} suggest the existence of a reduced higher genus Gromov--Witten theory formed by replacing contracted elliptic components with singularities\footnote{We have been informed by Jonathan Wise that his Ph.D. candidate Sebastian Bozlee has constructed these higher genus reduced Gromov--Witten invariants. A manuscript is in preparation.}. The discussion here is likely to carry over to this setting.

@@

We have produced a factorisation
\begin{equation*}\widetilde\Dcal\to\widetilde \Ecal\to\Ecal,
\end{equation*}
where the first map is flat and the second map is a logarithmic modification. $\widetilde\Dcal$ is a compactification of a torus bundle over $\widetilde\Ecal$ whose strata can be read off from the tropicalization; in particular, we have the information of how (polynomials in) the classes of boundary strata push forward to $\widetilde\Ecal$. In fact, we can say more and describe the generic fibre of $\widetilde\Dcal \to \widetilde\Ecal$:
\begin{lemma}\label{lem:generic_proj_bundle}
Generically over $\Ecal$, $\widetilde \Dcal\to\widetilde\Ecal$ is the projectivisation of a tautological vector bundle.
\end{lemma}
